### 2025-05-09 paper/paper.qmd

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Emily Su
thanks: "Code and data are available at: [https://github.com/moonsdust/data-quality](https://github.com/moonsdust/data-quality)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

\newpage

# References

### 2025-05-09 scripts/02-download_data.py
Updated installation instructions

@@ -4,7 +4,7 @@
 # Date: 7 May 2025
 # Contact: em.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Have the required packages ('requests', 'csv', and 'io') installed.
+# Pre-requisites: Have the required packages installed by following the instructions in README.md under "Setup".
 
 
 #### Workspace setup ####

### 2025-05-09 scripts/03-clean_data.py
Cleaning script with selected columns made (WIP)

@@ -1,0 +1,37 @@
+#### Preamble ####
+# Purpose: Cleans the raw data from Open Data Toronto
+# Author: Emily Su
+# Date: 9 May 2025
+# Contact: em.su@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Have ran 02-download_data.py and have installed the required packages mentioned in 
+# Workspace setup by running requirements.txt.
+
+#### Workspace setup ####
+import polars as pl
+import datetime
+
+#### Clean data ####
+# Read in CSV file
+the_raw_data = pl.read_csv("data/01-raw_data/raw_data.csv")
+
+# Select specific columns
+selected_columns = ["package", "accessibility", "completeness", "freshness", "metadata", "usability", "score", "grade", "recorded_at"]
+
+selected_df = the_raw_data.select(selected_columns)
+
+# # Filter to only rows that have data
+# filtered_df = selected_df.filter(the_raw_data["OCCUPIED_BEDS"].is_not_null())
+
+# print(filtered_df.head())
+
+# renamed_df = filtered_df.rename({"OCCUPANCY_DATE": "date",
+#                                  "SHELTER_ID": "Shelters",
+#                                  "CAPACITY_ACTUAL_BED": "Capacity",
+#                                  "OCCUPIED_BEDS": "Usage"
+#                                  })
+
+# # print(renamed_df.head())
+
+# #### Save data ####
+# renamed_df.write_csv("outputs/data/analysis_data.csv")

### 2025-05-09 scripts/03-clean_data.R
Delete scripts/03-clean_data.R

@@ -1,44 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Emily Su
-# Date: 9 May 2025
-# Contact: em.su@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-#### Workspace setup ####
-library(tidyverse)
-
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
-
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
-
-#### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")

### 2025-05-13 paper/paper.qmd
Added structure to paper

@@
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
-format: pdf
+format:
+  pdf:
+    toc: true
 number-sections: true
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+# Results {#sec-results}
 
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-# Results
-
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-## Weaknesses and next steps
+## Areas of improvement
 
@@
 
-# Appendix {-}
-
-# Additional data details
+# Appendix {#sec-appendix}
 

### 2025-05-13 scripts/03-clean_data.py
Finished cleaning script

@@ -9,29 +9,18 @@
 
 #### Workspace setup ####
 import polars as pl
-import datetime
 
 #### Clean data ####
 # Read in CSV file
 the_raw_data = pl.read_csv("data/01-raw_data/raw_data.csv")
 
 # Select specific columns
-selected_columns = ["package", "accessibility", "completeness", "freshness", "metadata", "usability", "score", "grade", "recorded_at"]
+selected_columns = ["package", "accessibility", "completeness", "freshness", "metadata", "usability", "grade"]
 
 selected_df = the_raw_data.select(selected_columns)
 
-# # Filter to only rows that have data
-# filtered_df = selected_df.filter(the_raw_data["OCCUPIED_BEDS"].is_not_null())
+# Filter to only rows that have no NONE data
+filtered_df = selected_df.filter(the_raw_data["package"].is_not_null() & the_raw_data["accessibility"].is_not_null() & the_raw_data["completeness"].is_not_null() & the_raw_data["freshness"].is_not_null() & the_raw_data["metadata"].is_not_null() & the_raw_data["usability"].is_not_null() & the_raw_data["grade"].is_not_null())
 
-# print(filtered_df.head())
-
-# renamed_df = filtered_df.rename({"OCCUPANCY_DATE": "date",
-#                                  "SHELTER_ID": "Shelters",
-#                                  "CAPACITY_ACTUAL_BED": "Capacity",
-#                                  "OCCUPIED_BEDS": "Usage"
-#                                  })
-
-# # print(renamed_df.head())
-
-# #### Save data ####
-# renamed_df.write_csv("outputs/data/analysis_data.csv")
+#### Save data ####
+filtered_df.write_csv("data/02-analysis_data/analysis_data.csv")

### 2025-05-13 scripts/03-clean_data.py
Updated analysis dataset

@@ -15,12 +15,12 @@ import polars as pl
 the_raw_data = pl.read_csv("data/01-raw_data/raw_data.csv")
 
 # Select specific columns
-selected_columns = ["package", "accessibility", "completeness", "freshness", "metadata", "usability", "grade"]
+selected_columns = ["package", "completeness", "freshness", "metadata", "usability", "grade"]
 
 selected_df = the_raw_data.select(selected_columns)
 
 # Filter to only rows that have no NONE data
-filtered_df = selected_df.filter(the_raw_data["package"].is_not_null() & the_raw_data["accessibility"].is_not_null() & the_raw_data["completeness"].is_not_null() & the_raw_data["freshness"].is_not_null() & the_raw_data["metadata"].is_not_null() & the_raw_data["usability"].is_not_null() & the_raw_data["grade"].is_not_null())
+filtered_df = selected_df.filter(the_raw_data["package"].is_not_null() & the_raw_data["completeness"].is_not_null() & the_raw_data["freshness"].is_not_null() & the_raw_data["metadata"].is_not_null() & the_raw_data["usability"].is_not_null() & the_raw_data["grade"].is_not_null())
 
 #### Save data ####
 filtered_df.write_csv("data/02-analysis_data/analysis_data.csv")

### 2025-05-13 scripts/03-clean_data.py
Updated to python 3.10 and updated analysis data

@@ -15,12 +15,12 @@ import polars as pl
 the_raw_data = pl.read_csv("data/01-raw_data/raw_data.csv")
 
 # Select specific columns
-selected_columns = ["package", "completeness", "freshness", "metadata", "usability", "grade"]
+selected_columns = ["package", "accessibility", "completeness", "freshness", "metadata", "usability", "grade"]
 
 selected_df = the_raw_data.select(selected_columns)
 
 # Filter to only rows that have no NONE data
-filtered_df = selected_df.filter(the_raw_data["package"].is_not_null() & the_raw_data["completeness"].is_not_null() & the_raw_data["freshness"].is_not_null() & the_raw_data["metadata"].is_not_null() & the_raw_data["usability"].is_not_null() & the_raw_data["grade"].is_not_null())
+filtered_df = selected_df.filter(the_raw_data["package"].is_not_null() & the_raw_data["accessibility"].is_not_null() & the_raw_data["completeness"].is_not_null() & the_raw_data["freshness"].is_not_null() & the_raw_data["metadata"].is_not_null() & the_raw_data["usability"].is_not_null() & the_raw_data["grade"].is_not_null())
 
 #### Save data ####
 filtered_df.write_csv("data/02-analysis_data/analysis_data.csv")

### 2025-05-13 scripts/04-test_analysis_data.R
removed R script for test analysis

@@ -1,69 +1,0 @@
-#### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Emily Su
-# Date: 9 May 2025
-# Contact: em.su@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(testthat)
-
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
-
-
-#### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
-})
-
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
-})
-
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
-})
-
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
-})
-
-# Test that the 'state' column is character type
-test_that("'state' is character", {
-  expect_type(analysis_data$state, "character")
-})
-
-# Test that there are no missing values in the dataset
-test_that("no missing values in dataset", {
-  expect_true(all(!is.na(analysis_data)))
-})
-
-# Test that 'division' contains unique values (no duplicates)
-test_that("'division' column contains unique values", {
-  expect_equal(length(unique(analysis_data$division)), 151)
-})
-
-# Test that 'state' contains only valid Australian state or territory names
-valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", "Western Australia", 
-                  "Tasmania", "Northern Territory", "Australian Capital Territory")
-test_that("'state' contains valid Australian state names", {
-  expect_true(all(analysis_data$state %in% valid_states))
-})
-
-# Test that there are no empty strings in 'division', 'party', or 'state' columns
-test_that("no empty strings in 'division', 'party', or 'state' columns", {
-  expect_false(any(analysis_data$division == "" | analysis_data$party == "" | analysis_data$state == ""))
-})
-
-# Test that the 'party' column contains at least 2 unique values
-test_that("'party' column contains at least 2 unique values", {
-  expect_true(length(unique(analysis_data$party)) >= 2)
-})

### 2025-05-13 scripts/04-test_analysis_data.py
Finished testing of analysis data

@@ -1,0 +1,96 @@
+#### Preamble ####
+# Purpose: Tests the cleaned dataset from Open Data Toronto 
+# Author: Emily Su
+# Date: 9 May 2025
+# Contact: em.su@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Have ran 02-download_data.py and 03-clean_data.py and have installed the required packages mentioned in 
+# Workspace setup by running requirements.txt.
+
+
+#### Workspace setup ####
+import polars as pl
+import pointblank as pb
+
+# Reference https://posit.co/blog/introducing-pointblank-for-python/
+# Read in CSV file
+cleaned_data = pl.read_csv("data/02-analysis_data/analysis_data.csv")
+
+#### Test data ####
+# Check columns type
+schema = pb.Schema(
+    columns=[
+        ("package", "String"),
+        ("accessibility", "Int64"),
+        ("completeness", "Float64"),
+        ("freshness", "Float64"),
+        ("metadata", "Float64"),
+        ("usability", "Float64"),
+        ("grade", "String"),
+    ]
+)
+check_columns_test = (
+    pb.Validate(
+        data=cleaned_data,
+        tbl_name="cleaned_data",
+        label="Check Columns Test"
+    )
+    .col_schema_match(schema=schema)
+    .interrogate()
+)
+
+# Check if all the columns don't have a NONE/NULL value
+check_for_null_test = (
+    pb.Validate(
+        data=cleaned_data,
+        tbl_name="cleaned_data",
+        label="Check NULL Test"
+    )
+     .col_vals_not_null(
+      columns=["package", "accessibility", "completeness", "freshness", "metadata", "usability", "grade"]
+   )
+    .interrogate()
+)
+
+# None of the columns in string type columns are ""
+check_for_empty_strings_test = (
+    pb.Validate(
+        data=cleaned_data,
+        tbl_name="cleaned_data",
+        label="Check for empty strings test"
+    )
+    .col_vals_regex(
+        columns=["package", "grade"],# check that all string columns are non-empty strings
+        pattern=r"(.|\s)*\S(.|\s)*" # Referenced: https://posit-dev.github.io/pointblank/demos/column-selector-functions/index.html
+    )
+    .interrogate()
+)
+
+# Check min and max for all numeric type columns except usability
+check_min_max_test = (
+    pb.Validate(
+        data=cleaned_data,
+        tbl_name="cleaned_data",
[…]

### 2025-05-14 scripts/03-clean_data.py
Updated analysis dataset

@@ -15,12 +15,12 @@ import polars as pl
 the_raw_data = pl.read_csv("data/01-raw_data/raw_data.csv")
 
 # Select specific columns
-selected_columns = ["package", "accessibility", "completeness", "freshness", "metadata", "usability", "grade"]
+selected_columns = ["accessibility", "completeness", "freshness", "metadata", "usability", "grade"]
 
 selected_df = the_raw_data.select(selected_columns)
 
 # Filter to only rows that have no NONE data
-filtered_df = selected_df.filter(the_raw_data["package"].is_not_null() & the_raw_data["accessibility"].is_not_null() & the_raw_data["completeness"].is_not_null() & the_raw_data["freshness"].is_not_null() & the_raw_data["metadata"].is_not_null() & the_raw_data["usability"].is_not_null() & the_raw_data["grade"].is_not_null())
+filtered_df = selected_df.filter(the_raw_data["accessibility"].is_not_null() & the_raw_data["completeness"].is_not_null() & the_raw_data["freshness"].is_not_null() & the_raw_data["metadata"].is_not_null() & the_raw_data["usability"].is_not_null() & the_raw_data["grade"].is_not_null())
 
 #### Save data ####
 filtered_df.write_csv("data/02-analysis_data/analysis_data.csv")

### 2025-05-14 scripts/04-test_analysis_data.py

@@ -20,7 +20,6 @@ cleaned_data = pl.read_csv("data/02-analysis_data/analysis_data.csv")
 # Check columns type
 schema = pb.Schema(
     columns=[
-        ("package", "String"),
         ("accessibility", "Int64"),
         ("completeness", "Float64"),
         ("freshness", "Float64"),
@@ -47,7 +46,7 @@ check_for_null_test = (
         label="Check NULL Test"
     )
      .col_vals_not_null(
-      columns=["package", "accessibility", "completeness", "freshness", "metadata", "usability", "grade"]
+      columns=["accessibility", "completeness", "freshness", "metadata", "usability", "grade"]
    )
     .interrogate()
 )
@@ -60,7 +59,7 @@ check_for_empty_strings_test = (
         label="Check for empty strings test"
     )
     .col_vals_regex(
-        columns=["package", "grade"],# check that all string columns are non-empty strings
+        columns=["grade"],# check that all string columns are non-empty strings
         pattern=r"(.|\s)*\S(.|\s)*" # Referenced: https://posit-dev.github.io/pointblank/demos/column-selector-functions/index.html
     )
     .interrogate()

### 2025-05-14 scripts/00-simulate_data.py
Added scripts and simulated dataset

@@ -1,8 +1,7 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  # state and party that won each division.
+# Purpose: Simulates a dataset of the data quality of datasets on Open Data Toronto
 # Author: Emily Su
-# Date: 9 May 2025
+# Date: 13 May 2025
 # Contact: em.su@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
@@ -13,36 +12,33 @@
 #### Workspace setup ####
 import polars as pl
 import numpy as np
-np.random.seed(853)
+np.random.seed(646)
 
 
 #### Simulate data ####
-# State names
-states = [
-    "New South Wales", "Victoria", "Queensland", "South Australia", 
-    "Western Australia", "Tasmania", "Northern Territory", 
-    "Australian Capital Territory"
+# Grade
+grade = [
+  "Bronze", "Silver", "Gold"
 ]
 
-# Political parties
-parties = ["Labor", "Liberal", "Greens", "National", "Other"]
-
-# Probabilities for state and party distribution
-state_probs = [0.25, 0.25, 0.15, 0.1, 0.1, 0.1, 0.025, 0.025]
-party_probs = [0.40, 0.40, 0.05, 0.1, 0.05]
-
 # Generate the data using numpy and polars
-divisions = [f"Division {i}" for i in range(1, 152)]
-states_sampled = np.random.choice(states, size=151, replace=True, p=state_probs)
-parties_sampled = np.random.choice(parties, size=151, replace=True, p=party_probs)
+accessibility = np.random.randint(low=0, high=1, size=151, dtype=int)
+completeness = np.random.random()
+freshness = np.random.random()
+metadata = np.random.random()
+usability = np.random.random()
+grade = np.random.choice(grade, size=151, replace=True, p=[0.5, 0.25, 0.25])
+
 
 # Create a polars DataFrame
 analysis_data = pl.DataFrame({
-    "division": divisions,
-    "state": states_sampled,
-    "party": parties_sampled
+    "accessibility": accessibility,
+    "completeness": completeness,
+    "freshness": freshness,
+    "metadata": metadata,
+    "usability": usability,
+    "grade": grade
 })
-
 
 #### Save data ####
 analysis_data.write_csv("data/00-simulated_data/simulated_data.csv")

### 2025-05-14 scripts/01-test_simulated_data.py

@@ -1,0 +1,95 @@
+#### Preamble ####
+# Purpose: Tests the simulated dataset
+# Author: Emily Su
+# Date: 13 May 2025
+# Contact: em.su@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Have ran 00-simulate_data.py and have installed the required packages mentioned in 
+# Workspace setup by running requirements.txt.
+
+
+#### Workspace setup ####
+import polars as pl
+import pointblank as pb
+
+# Reference https://posit.co/blog/introducing-pointblank-for-python/
+# Read in CSV file
+simulated_data = pl.read_csv("data/00-simulated_data/simulated_data.csv")
+
+#### Test data ####
+# Check columns type
+schema = pb.Schema(
+    columns=[
+        ("accessibility", "Int64"),
+        ("completeness", "Float64"),
+        ("freshness", "Float64"),
+        ("metadata", "Float64"),
+        ("usability", "Float64"),
+        ("grade", "String"),
+    ]
+)
+check_columns_test = (
+    pb.Validate(
+        data=simulated_data,
+        tbl_name="cleaned_data",
+        label="Check Columns Test"
+    )
+    .col_schema_match(schema=schema)
+    .interrogate()
+)
+
+# Check if all the columns don't have a NONE/NULL value
+check_for_null_test = (
+    pb.Validate(
+        data=simulated_data,
+        tbl_name="cleaned_data",
+        label="Check NULL Test"
+    )
+     .col_vals_not_null(
+      columns=["accessibility", "completeness", "freshness", "metadata", "usability", "grade"]
+   )
+    .interrogate()
+)
+
+# None of the columns in string type columns are ""
+check_for_empty_strings_test = (
+    pb.Validate(
+        data=simulated_data,
+        tbl_name="cleaned_data",
+        label="Check for empty strings test"
+    )
+    .col_vals_regex(
+        columns=["grade"],# check that all string columns are non-empty strings
+        pattern=r"(.|\s)*\S(.|\s)*" # Referenced: https://posit-dev.github.io/pointblank/demos/column-selector-functions/index.html
+    )
+    .interrogate()
+)
+
+# Check min and max for all numeric type columns except usability
+check_min_max_test = (
+    pb.Validate(
+        data=simulated_data,
+        tbl_name="cleaned_data",
+        label="Check Min/Max Test"
+    )
[…]

### 2025-05-14 paper/paper.qmd
Made changes to paper

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Accessible and Usable Datasets yet Majority are Bronze-Graded and not Updated Regularly"
+subtitle: "An analysis of the data quality of datasets available on the Open Data Toronto Portal (As of May 13, 2025)"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "As one of the central hubs for data on the City of Toronto that are used in the media and in policy-making, we conducted analysis on the quality of Open Data Toronto's data catalogue.
+We found that despite Open Data Toronto's extensive dataset catalogue being accessible and usable, 56% of their datasets are graded bronze and bronze-graded datasets are less likely to be updated the more completed they are.
+However we also found that metadata fields were less likely to be filled for gold-graded and silver-graded datasets.
+These findings can help reporters, policy-makers, and anyone interested in using datasets from Open Data Toronto's catalogue make an inform decision of what datasets to choose."
 format:
@@
 bibliography: references.bib
+execute:
+  python: ".venv/bin/python"
+  external: true
 ---
-
-```{r}
+```{python}
 #| include: false
@@
 
-library(tidyverse)
-library(palmerpenguins)
+import polars as pl
+import matplotlib.pyplot as plt
+import seaborn as sns
+import pandas as pd
+import pyarrow
+
+# Import dataset 
+cleaned_data = pl.read_csv("../data/02-analysis_data/analysis_data.csv")
+cleaned_data = cleaned_data.to_pandas()
 ```
@@
 
-Overview paragraph
+In the analysis of our paper, we looked at the following questions: What is the quality of the datasets on Open Data Toronto? What are the features of different types of datasets on Open Data Toronto?
 
-Estimand paragraph
-
-Results paragraph
-
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+For the remainder of the paper, the data section (@sec-data) looks at the data used and how it was retrieved alongside the characteristics of the data and our variables of interest for our analysis.
+In the results section (@sec-results) we looked at the data more in-depth through graphs.
+Finally, the appendix (@sec-appendix) includes acknowledgements and any additional information related to the paper.
 
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+The dataset used in the paper comes from Open Data Toronto portal titled "Catalogue quality scores" [@citeOpenDataToronto].
+Other datasets like "Toronto Open Data Intake" were considered in the analysis of the paper however, it does not indicate the quality of the datasets that are being requested.
+This specific dataset looks at the quality of the datasets available from the Open Data Toronto catalogue to inform others
+how valuable certain datasets to be used for various situations like civic issues.
+The datasets are scored based on characteristics such as its accessibility, completeness, freshness, metadata, and usability, which are then calculated together to
+give a dataset a grade that is displayed alongside a trophy icon under the details section on a dataset page on Open Data Toronto portal [@citeOpenDataToronto].
 
-Overview text
+We used the programming language Python [@citePython], the statistical programming language R [@citeR], and the following libraries to download, clean, analyze, and test the dataset and the overall paper itself: Requests [@citeRequests], datetime [@citeDatetime], Matplotlib [@citeMatplotlib], numpy [@citenumpy], pandas [@citepandas], Polars [@citepolars], Pydantic [@citePydantic], seaborn [@citeSeaborn], Pointblank [@citePointblank], and Pyarrow [@citePyarrow].
+
+We retrieved the raw dataset by calling the Open Data Toronto API [@citeOpenDataToronto] using the Requests library [@citeRequests] and downloading the file as a CSV.
+There are 39,580 total observations in the cleaned dataset with each observation being a dataset in the catalogue.
+The cleaned dataset look as follows @tbl-dataset-preview:
+
+```{python}
+#| label: tbl-dataset-preview
+#| tbl-cap: Preview of dataset on Open Data Toronto's Catalogue quality scores as of May 13, 2025
+#| echo: false
+#| warning: false
+#| message: false
+
+# Create table
+cleaned_data[["accessibility", "completeness", "freshness", "metadata", "usability", "grade"]].head(5)
+```
 
 ## Measurement
+The Information & Technology department at Open Data Toronto collected 
+[@citeCatalogue]
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
-
-## Outcome variables
+## Variables of Interest
 
@@
 
-# Discussion {#sec-discussion}
+## Grade of datasets 
+```{python}
+#| label: fig-grade-bar
+#| fig-cap: "Number of datasets on Open Data Toronto graded bronze, silver, and gold as of May 13, 2025"
+#| warning: false
+#| message: false
+#| echo: false
 
-## First discussion point {#sec-first-point}
+# Set styling
+sns.set_theme(style="whitegrid")
+plt.figure(figsize=(12, 6))
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+# Code from: https://stackoverflow.com/questions/49044131/how-to-add-data-labels-to-seaborn-countplot-factorplot
+# Create countplot
+ax = sns.countplot(cleaned_data, x="grade",
+                   order=cleaned_data["grade"].value_counts(ascending=False).index);
+# Create label
+lbls = [f'{p[0]} ({p[1]:.0f}%)' for p in zip(cleaned_data['grade'].value_counts(ascending=False), cleaned_data['grade'].value_counts(ascending=False, normalize=True).values * 100)]
+# Combine countplot and labels into bar_label
+ax.bar_label(container=ax.containers[0], labels=lbls)
 
-## Second discussion point
+# Add labels and title
+plt.xlabel("Grade of the Dataset")
+plt.ylabel("Number of Datasets")
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+# Adjust layout to prevent clipping of tick-labels
+plt.tight_layout()
 
-## Third discussion point
+# Display the plot
+plt.show()
+```
 
-## Areas of improvement
+As of May 13, 2025, @fig-grade-bar shows that 56% of datasets on the Open Data Toronto portal had a grade of "bronze".
+Following this, 25% of datasets are graded "gold" and finally 19% of datasets are graded "silver".
+This means half of the datasets on the Open Data Toronto portal are ranked "bronze".
 
-Weaknesses and next steps should also be included.
+## Accessibility scores of datasets
+```{python}
+#| label: fig-accessibility-bar
+#| fig-cap: "Number of datasets with accessiblity score of 1 on Open Data Toronto as of May 13, 2025"
+#| warning: false
+#| message: false
+#| echo: false
+
+# Set styling
+sns.set_theme(style="whitegrid")
+plt.figure(figsize=(12, 6))
+
+# Code from: https://stackoverflow.com/questions/49044131/how-to-add-data-labels-to-seaborn-countplot-factorplot
+
+# filtered data for accessiblity score of 1 
+filtered_data = cleaned_data[cleaned_data["accessibility"] == 1]
+
+# Create countplot
+ax = sns.countplot(data = filtered_data, x="grade", order=cleaned_data["grade"].value_counts(ascending=False).index);
+
+# Create label
+lbls = [f'{p[0]}' for p in zip(filtered_data['grade'].value_counts(ascending=False))]
+# Combine countplot and labels into bar_label
+ax.bar_label(container=ax.containers[0], labels=lbls)
+
+# Add labels and title
+plt.xlabel("Grade of the Dataset")
+plt.ylabel("Number of datasets")
+
+# Adjust layout to prevent clipping of tick-labels
+plt.tight_layout()
+
+# Display the plot
+plt.show()
+```
+
+As of May 13, 2025, @fig-accessibility-bar indicates 22294 bronze-graded datasets on the Open Data Toronto portal scored 1 for accessiblity.
+For gold-graded datasets, it was 9698 datasets and for silver-graded datasets, it was 7588 datasets.
+
+## The relationship between completeness and usability scores of datasets 
+```{python}
+#| label: fig-completeness-usability-scatterplot
+#| fig-cap: "The relationship between completeness scores and usability scores of datasets on Open Data Toronto across different grades as of May 13, 2025"
+#| warning: false
+#| message: false
+#| echo: false
+
+# Set styling
+sns.set_theme(style="whitegrid")
+
+# Create regplot
+ax = sns.lmplot(data = cleaned_data, x = "completeness", y = "usability", hue="grade", scatter_kws={'alpha': 0.3})
+sns.move_legend(ax, "upper left", bbox_to_anchor=(1, 1))
+
+# Add labels and title
+plt.xlabel("Completeness of the Dataset")
+plt.ylabel("Usability of the Dataset")
+
+# Adjust layout to prevent clipping of tick-labels
+plt.tight_layout()
+
+# Display the plot
+plt.show()
+```
+
+As seen in @fig-completeness-usability-scatterplot, for all grades, there's a slight positive relationship between the completeness of a dataset on the Open Data Toronto portal and its usability.
+However, this relationship is more apparent with the datasets that are graded bronze.
+This means as the completeness score increases, the usability score of the dataset increases.
+
+## Completeness versus metadata completion scores of datasets
+```{python}
+#| label: fig-completeness-metadata-scatterplot
+#| fig-cap: "The relationship between completeness scores and metadata completion scores of datasets on Open Data Toronto across different grades as of May 13, 2025"
+#| warning: false
+#| message: false
+#| echo: false
+
+# Set styling
+sns.set_theme(style="whitegrid")
+
+# Create regplot
+ax = sns.lmplot(data = cleaned_data, x = "completeness", y = "metadata", hue="grade", scatter_kws={'alpha': 0.3})
+sns.move_legend(ax, "upper left", bbox_to_anchor=(1, 1))
+
+# Add labels and title
+plt.xlabel("Completeness of the Dataset")
+plt.ylabel("Metadata Completion Score of the Dataset")
+
+# Adjust layout to prevent clipping of tick-labels
+plt.tight_layout()
+
+# Display the plot
+plt.show()
+```
+
+In term of the relationship between the completeness score and the metadata completion score of datasets on the Open Data Toronto portal in @fig-completeness-metadata-scatterplot, there is a slight positive relationship for bronze-graded datasets while there is a negative relationship for silver-graded and gold-graded datasets.
+This means that for bronze-graded datasets, as the completeness score increasing, the metadata score increases.
+However for gold-graded and silver-graded datasets, as the completeness score increases, the metadata score decreases.
+
+## Completeness versus freshness scores of datasets 
+```{python}
+#| label: fig-completeness-freshness-scatterplot
+#| fig-cap: "The relationship between completeness scores and freshness scores of datasets on Open Data Toronto across different grades as of May 13, 2025"
+#| warning: false
+#| message: false
+#| echo: false
+
+# Set styling
+sns.set_theme(style="whitegrid")
+
+# Create regplot
+ax = sns.lmplot(data = cleaned_data, x = "completeness", y = "freshness", hue="grade", scatter_kws={'alpha': 0.3})
+sns.move_legend(ax, "upper left", bbox_to_anchor=(1, 1))
+
+# Add labels and title
+plt.xlabel("Completeness of the Dataset")
+plt.ylabel("Freshness of the Dataset")
+
+# Adjust layout to prevent clipping of tick-labels
+plt.tight_layout()
+
+# Display the plot
+plt.show()
+```
+
+As of May 13, 2025, when we look at the relationship between completeness score and freshness score in @fig-completeness-freshness-scatterplot, there is a negative relationship for bronze-graded datasets, positive relationship for gold-graded datasets, and there is almost no relationship between the two for silver-graded datasets.
+This means that for bronze-graded datasets as the completeness score increases, the freshness score decreases while for gold-graded datasets as the completeness score increases, the freshness score increases.
+However with silver-graded datasets, the freshness score doesn't change as the completion score increases.
 
@@
 
+## Acknowledgments
+
+We would like to thank @tellingstories for providing assistance with the code used to produce the graphs in this paper.
+
 \newpage

### 2025-05-14 scripts/01-test_simulated_data.R
Delete scripts/01-test_simulated_data.R

@@ -1,89 +1,0 @@
-#### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Emily Su
-# Date: 9 May 2025
-# Contact: em.su@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
-
-#### Workspace setup ####
-library(tidyverse)
-
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
-
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
-
-
-#### Test data ####
-
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
-} else {
-  stop("Test Failed: The dataset does not have 151 rows.")
-}
-
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
-} else {
-  stop("Test Failed: The dataset does not have 3 columns.")
-}
-
-# Check if all values in the 'division' column are unique
-if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
-  message("Test Passed: All values in 'division' are unique.")
-} else {
-  stop("Test Failed: The 'division' column contains duplicate values.")
-}
-
-# Check if the 'state' column contains only valid Australian state names
-valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", 
-                  "Western Australia", "Tasmania", "Northern Territory", 
-                  "Australian Capital Territory")
-
-if (all(analysis_data$state %in% valid_states)) {
-  message("Test Passed: The 'state' column contains only valid Australian state names.")
-} else {
-  stop("Test Failed: The 'state' column contains invalid state names.")
-}
-
-# Check if the 'party' column contains only valid party names
-valid_parties <- c("Labor", "Liberal", "Greens", "National", "Other")
-
-if (all(analysis_data$party %in% valid_parties)) {
-  message("Test Passed: The 'party' column contains only valid party names.")
-} else {
-  stop("Test Failed: The 'party' column contains invalid party names.")
-}
-
[…]

### 2025-05-14 paper/paper.qmd
Updated paper

@@
 
-In the analysis of our paper, we looked at the following questions: What is the quality of the datasets on Open Data Toronto? What are the features of different types of datasets on Open Data Toronto?
+In the analysis of our paper, we looked at the following questions: What is the quality of the datasets on the Open Data Toronto portal?
+What are the features of different types of datasets on Open Data Toronto?
 
@@
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+Our variables of interest are "accessibility", "completeness", "freshness", "metadata", "usability", and "grade".
 

### 2025-05-14 scripts/03-clean_data.py
Put checks on one line each in cleaned data script

@@ -20,7 +20,12 @@ selected_columns = ["accessibility", "completeness", "freshness", "metadata", "u
 selected_df = the_raw_data.select(selected_columns)
 
 # Filter to only rows that have no NONE data
-filtered_df = selected_df.filter(the_raw_data["accessibility"].is_not_null() & the_raw_data["completeness"].is_not_null() & the_raw_data["freshness"].is_not_null() & the_raw_data["metadata"].is_not_null() & the_raw_data["usability"].is_not_null() & the_raw_data["grade"].is_not_null())
+filtered_df = selected_df.filter(the_raw_data["accessibility"].is_not_null() & 
+                                 the_raw_data["completeness"].is_not_null() & 
+                                 the_raw_data["freshness"].is_not_null() & 
+                                 the_raw_data["metadata"].is_not_null() & 
+                                 the_raw_data["usability"].is_not_null() & 
+                                 the_raw_data["grade"].is_not_null())
 
 #### Save data ####
 filtered_df.write_csv("data/02-analysis_data/analysis_data.csv")

### 2025-05-14 paper/paper.qmd
Updated abstract in qmd

@@
 date-format: long
-abstract: "As one of the central hubs for data on the City of Toronto that are used in the media and in policy-making, we conducted analysis on the quality of Open Data Toronto's data catalogue.
-We found that despite Open Data Toronto's extensive dataset catalogue being accessible and usable, 56% of their datasets are graded bronze and bronze-graded datasets are less likely to be updated the more completed they are.
-However we also found that metadata fields were less likely to be filled for gold-graded and silver-graded datasets.
-These findings can help reporters, policy-makers, and anyone interested in using datasets from Open Data Toronto's catalogue make an inform decision of what datasets to choose."
+abstract: "As one of the central hubs for Toronto-related data, we conducted analysis on the data quality of Open Data Toronto's catalogue.
+We found that despite Open Data Toronto's extensive dataset catalogue being accessible and usable, 56% of their datasets are graded bronze and bronze-graded datasets are less likely to be updated the more complete they are.
+However, we also found that metadata fields were less likely to be filled for gold-graded and silver-graded datasets.
+These findings can help reporters, policy-makers, and anyone interested in using datasets from Open Data Toronto's catalogue make an inform decision of what datasets to choose for purposes like journalism and in policy-making."
 format:

### 2025-05-15 paper/paper.qmd
Revamped a lot of the paper and updated README.md

@@
 ---
-title: "Accessible and Usable Datasets yet Majority are Bronze-Graded and not Updated Regularly"
+title: "Accessible and Completed Datasets yet Majority are Bronze-Graded and not Updated Regularly with Missing Metadata"
 subtitle: "An analysis of the data quality of datasets available on the Open Data Toronto Portal (As of May 13, 2025)"
@@
 abstract: "As one of the central hubs for Toronto-related data, we conducted analysis on the data quality of Open Data Toronto's catalogue.
-We found that despite Open Data Toronto's extensive dataset catalogue being accessible and usable, 56% of their datasets are graded bronze and bronze-graded datasets are less likely to be updated the more complete they are.
-However, we also found that metadata fields were less likely to be filled for gold-graded and silver-graded datasets.
-These findings can help reporters, policy-makers, and anyone interested in using datasets from Open Data Toronto's catalogue make an inform decision of what datasets to choose for purposes like journalism and in policy-making."
+We found that despite Open Data Toronto's extensive dataset catalogue being accessible and having minimal missing data, 56% of their datasets are graded bronze and bronze-graded datasets are less likely to be updated and have completed metadata fields.
+These findings can help raise awareness to Open Data Toronto whose datasets play an important role in news reporting and policymaking and anyone who is interested using datasets from Open Data Toronto's catalogue understand what goes behind the grade given to datasets."
 format:
@@
 
-For the remainder of the paper, the data section (@sec-data) looks at the data used and how it was retrieved alongside the characteristics of the data and our variables of interest for our analysis.
+For the remainder of the paper, the data section (@sec-data) looks at the data used and how it was retrieved alongside the characteristics of the data, the data's limitations, and our variables of interest for our analysis.
 In the results section (@sec-results) we looked at the data more in-depth through graphs.
@@
 This specific dataset looks at the quality of the datasets available from the Open Data Toronto catalogue to inform others
-how valuable certain datasets to be used for various situations like civic issues.
+how valuable certain datasets are to be used for various situations like reporting on civic issues.
 The datasets are scored based on characteristics such as its accessibility, completeness, freshness, metadata, and usability, which are then calculated together to
@@
 There are 39,580 total observations in the cleaned dataset with each observation being a dataset in the catalogue.
-The cleaned dataset look as follows @tbl-dataset-preview:
+@tbl-dataset-preview shows a preview of what the cleaned dataset looks like:
 
@@
 
+@tbl-summary-statistics shows the summary statistics of the cleaned dataset:
+
+```{python}
+#| label: tbl-summary-statistics
+#| tbl-cap: Summary statistics of dataset on Open Data Toronto's Catalogue quality scores as of May 13, 2025
+#| echo: false
+#| warning: false
+#| message: false
+
+# Create table
+cleaned_data.describe()
+```
+
 ## Measurement
-The Information & Technology department at Open Data Toronto collected 
-[@citeCatalogue]
+The Information & Technology department at Open Data Toronto collected datasets on their portal using the CKAN API
+[@citeCatalogue; @citeCatalogueSteps]
+
+The score for each dataset is recalculated every week [@citeCatalogue]
 
@@
 
-## Grade of datasets 
+## Grade and accessibility of datasets 
 ```{python}
-#| label: fig-grade-bar
-#| fig-cap: "Number of datasets on Open Data Toronto graded bronze, silver, and gold as of May 13, 2025"
+#| label: fig-grade-accessibility-bar
+#| fig-cap: "Number of datasets and their accessibility on Open Data Toronto graded bronze, silver, and gold as of May 13, 2025"
 #| warning: false
@@
 # Code from: https://stackoverflow.com/questions/49044131/how-to-add-data-labels-to-seaborn-countplot-factorplot
+
 # Create countplot
 ax = sns.countplot(cleaned_data, x="grade",
-                   order=cleaned_data["grade"].value_counts(ascending=False).index);
+                   order=cleaned_data["grade"].value_counts(ascending=False).index, palette = "crest", hue = "accessibility");
 # Create label
@@
 
-As of May 13, 2025, @fig-grade-bar shows that 56% of datasets on the Open Data Toronto portal had a grade of "bronze".
+As of May 13, 2025, @fig-grade-accessibility-bar shows that 56% of datasets on the Open Data Toronto portal had a grade of "bronze".
 Following this, 25% of datasets are graded "gold" and finally 19% of datasets are graded "silver".
 This means half of the datasets on the Open Data Toronto portal are ranked "bronze".
-
-## Accessibility scores of datasets
-```{python}
-#| label: fig-accessibility-bar
-#| fig-cap: "Number of datasets with accessiblity score of 1 on Open Data Toronto as of May 13, 2025"
-#| warning: false
-#| message: false
-#| echo: false
-
-# Set styling
-sns.set_theme(style="whitegrid")
-plt.figure(figsize=(12, 6))
-
-# Code from: https://stackoverflow.com/questions/49044131/how-to-add-data-labels-to-seaborn-countplot-factorplot
-
-# filtered data for accessiblity score of 1 
-filtered_data = cleaned_data[cleaned_data["accessibility"] == 1]
-
-# Create countplot
-ax = sns.countplot(data = filtered_data, x="grade", order=cleaned_data["grade"].value_counts(ascending=False).index);
-
-# Create label
-lbls = [f'{p[0]}' for p in zip(filtered_data['grade'].value_counts(ascending=False))]
-# Combine countplot and labels into bar_label
-ax.bar_label(container=ax.containers[0], labels=lbls)
-
-# Add labels and title
-plt.xlabel("Grade of the Dataset")
-plt.ylabel("Number of datasets")
-
-# Adjust layout to prevent clipping of tick-labels
-plt.tight_layout()
-
-# Display the plot
-plt.show()
-```
-
-As of May 13, 2025, @fig-accessibility-bar indicates 22294 bronze-graded datasets on the Open Data Toronto portal scored 1 for accessiblity.
-For gold-graded datasets, it was 9698 datasets and for silver-graded datasets, it was 7588 datasets.
+However since all the datasets have an accessibility score of 1, which indicates they are accessible, and the mean accessibility score is 1 by @tbl-summary-statistics as well, it indicates all datasets on the Open Data Toronto portal can be accessed directly using methods like an API, tags or keywords, or automated data pipelines accessing Open Data Toronto's catalogue.
 
@@
 #| label: fig-completeness-usability-scatterplot
-#| fig-cap: "The relationship between completeness scores and usability scores of datasets on Open Data Toronto across different grades as of May 13, 2025"
+#| fig-cap: "The relationship between completeness scores and usability scores of Open Data Toronto's datasets across different grades as of May 13, 2025"
 #| warning: false
@@
 
-# Create regplot
-ax = sns.lmplot(data = cleaned_data, x = "completeness", y = "usability", hue="grade", scatter_kws={'alpha': 0.3})
-sns.move_legend(ax, "upper left", bbox_to_anchor=(1, 1))
+# Create lmplot
+ax = sns.lmplot(data = cleaned_data, x = "completeness", y = "usability", col = "grade", hue="grade", palette="crest", ci=None,
+    height=4, scatter_kws={"s": 50, "alpha": 0.2})
 
 # Add labels and title
-plt.xlabel("Completeness of the Dataset")
-plt.ylabel("Usability of the Dataset")
+ax.set(xlabel="Completeness of the Dataset", ylabel="Usability of the Dataset")
 
@@
 This means as the completeness score increases, the usability score of the dataset increases.
+We can also see most of the scores for bronze-graded datasets are more spread out along the completeness score axis compared to gold-graded and silver-graded datasets.
+This indicates that more bronze-grade datasets contain more missing data than the other graded datasets.
+Despite this, @fig-completeness-distribution shows that the completeness score of datasets on Open Data Toronto skews left with their peaks being above 0.6 (60%), this indicates that across all grades, the datasets have minimal missing data.
+@tbl-summary-statistics also indicates that the mean values for completeness and usability scores across all datasets are 0.87 (87%) and 0.84 (84%), respectively.
 
-## Completeness versus metadata completion scores of datasets
 ```{python}
-#| label: fig-completeness-metadata-scatterplot
-#| fig-cap: "The relationship between completeness scores and metadata completion scores of datasets on Open Data Toronto across different grades as of May 13, 2025"
+#| label: fig-completeness-distribution
+#| fig-cap: "The distribution of completeness scores of Open Data Toronto's datasets across different grades as of May 13, 2025"
 #| warning: false
@@
 
-# Create regplot
-ax = sns.lmplot(data = cleaned_data, x = "completeness", y = "metadata", hue="grade", scatter_kws={'alpha': 0.3})
-sns.move_legend(ax, "upper left", bbox_to_anchor=(1, 1))
+# Create density plot
+ax = sns.kdeplot(
+   data=cleaned_data, x="completeness", hue="grade",
+   fill=True, common_norm=False, palette="crest",
+   alpha=.3, linewidth=1.3,
+)
 
 # Add labels and title
-plt.xlabel("Completeness of the Dataset")
-plt.ylabel("Metadata Completion Score of the Dataset")
+plt.xlabel("Completeness Score of the Dataset")
+plt.ylabel("Number of Datasets")
 
@@
 
-In term of the relationship between the completeness score and the metadata completion score of datasets on the Open Data Toronto portal in @fig-completeness-metadata-scatterplot, there is a slight positive relationship for bronze-graded datasets while there is a negative relationship for silver-graded and gold-graded datasets.
-This means that for bronze-graded datasets, as the completeness score increasing, the metadata score increases.
-However for gold-graded and silver-graded datasets, as the completeness score increases, the metadata score decreases.
-
-## Completeness versus freshness scores of datasets 
+## Metadata completeness scores of datasets
 ```{python}
-#| label: fig-completeness-freshness-scatterplot
-#| fig-cap: "The relationship between completeness scores and freshness scores of datasets on Open Data Toronto across different grades as of May 13, 2025"
+#| label: fig-metadata-distribution
+#| fig-cap: "The distribution of metadata completeness scores of Open Data Toronto's datasets across different grades as of May 13, 2025"
 #| warning: false
@@
 
-# Create regplot
-ax = sns.lmplot(data = cleaned_data, x = "completeness", y = "freshness", hue="grade", scatter_kws={'alpha': 0.3})
-sns.move_legend(ax, "upper left", bbox_to_anchor=(1, 1))
+# Create density plot
+ax = sns.kdeplot(
+   data=cleaned_data, x="metadata", hue="grade",
+   fill=True, common_norm=False, palette="crest",
+   alpha=.3, linewidth=1.3,
+)
 
 # Add labels and title
-plt.xlabel("Completeness of the Dataset")
-plt.ylabel("Freshness of the Dataset")
+plt.xlabel("Metadata Completeness Score of the Dataset")
+plt.ylabel("Number of Datasets")
 
@@
 
-As of May 13, 2025, when we look at the relationship between completeness score and freshness score in @fig-completeness-freshness-scatterplot, there is a negative relationship for bronze-graded datasets, positive relationship for gold-graded datasets, and there is almost no relationship between the two for silver-graded datasets.
-This means that for bronze-graded datasets as the completeness score increases, the freshness score decreases while for gold-graded datasets as the completeness score increases, the freshness score increases.
-However with silver-graded datasets, the freshness score doesn't change as the completion score increases.
+@fig-metadata-distribution shows that the metadata score for all datasets of Open Data Toronto's datasets has a multimodal distribution.
+However, the distribution of gold-graded datasets skew left overall.
+This means that most of the gold-graded datasets have metadata that is almost or is completed filled on the Open Data Toronto portal.
+On the other hand, the distribution of bronze-graded datasets overall skew right with its largest peak being below a metadata score of 0.5 or 50%.
+This indicates that the metadata fields for bronze-graded datasets are not sufficiently field or yet not filled out on the Open Data Toronto portal.
+@tbl-summary-statistics indicates that the mean metadata completeness score is 0.47 (47%) for all datasets on the portal.
+This means that the average metadata completeness score is below 50% for all datasets on the portal.
+
+## Freshness scores of datasets 
+```{python}
+#| label: fig-freshness-distribution
+#| fig-cap: "The distribution of freshness scores of Open Data Toronto's datasets across different grades as of May 13, 2025"
+#| warning: false
+#| message: false
+#| echo: false
+
+# Set styling
+sns.set_theme(style="whitegrid")
+
+# Create density plot
+ax = sns.kdeplot(
+   data=cleaned_data, x="freshness", hue="grade",
+   fill=True, common_norm=False, palette="crest",
+   alpha=.3, linewidth=1.3,
+)
+
+# Add labels and title
+plt.xlabel("Freshness Score of the Dataset")
+plt.ylabel("Number of Datasets")
+
+# Adjust layout to prevent clipping of tick-labels
+plt.tight_layout()
+
+# Display the plot
+plt.show()
+```
+
+As of May 13, 2025, @fig-freshness-distribution indicates that for gold-graded and silver-graded datasets, their distributions skews left and that the highest peaks of their distributions are around a freshness score of 1.0 or 100%.
+This indicates that the datasets that are gold-graded and silver-graded are frequently updated.
+However with bronze-graded datasets, its distribution skews right with its highest peak being around a freshness score 0.0 or 0%.
+This indicates that the datasets are not updated frequently or at all.
+@tbl-summary-statistics also shows that the mean freshness score is 0.56 (56%) across all datasets.
 

### 2025-05-15 paper/paper.qmd
Finished methdology of data section

@@
 The datasets are scored based on characteristics such as its accessibility, completeness, freshness, metadata, and usability, which are then calculated together to
-give a dataset a grade that is displayed alongside a trophy icon under the details section on a dataset page on Open Data Toronto portal [@citeOpenDataToronto].
+give a dataset a grade. This grade is displayed alongside a trophy icon under the details section on a dataset page on Open Data Toronto portal [@citeOpenDataToronto].
 
@@
 ## Measurement
-The Information & Technology department at Open Data Toronto collected datasets on their portal using the CKAN API
-[@citeCatalogue; @citeCatalogueSteps]
+The Information & Technology department at Open Data Toronto collected datasets on their portal using the CKAN Datastore API
+[@citeCatalogue; @citeCatalogueSteps].
+Open Data Toronto uses a metric called the "Data Quality Score" in order to give each dataset in their catalogue a grade indicating "bronze", "silver", or "gold", which can be seen through on the webpage for each dataset on Open Data Toronto portal website.
+In order to create the "Data Quality Score", they assembled the Data Quality Working Group, which consist of a diverse group of people from consumers of datasets to people who create datasets [@citeCatalogue; @citeCatalogueSteps].
+Open Data Toronto first reviewed various literature such as academic papers and industry white papers to compile 15 dimensions used to measure quality [@citeCatalogue; @citeCatalogueSteps].
+The dimensions selected were as follows: Interpretability ("How easy it is to understand the data?"), Usability ("How easy is it to work with the data?"), Metadata ("Is the data well described?"), Freshness ("How close to creation time is the data published?"), Granularity ("How atomic is the data?"), Completeness ("How much data is missing?"), and Accessibility ("Is the data easy to access?") [@citeCatalogue; @citeCatalogueSteps].
+From there the Data Quality Working Group were surveyed and asked to rank the importance of each dimension when it comes to assessing data quality where 1 represents the most important and 7 being the least important.
+The results from the survey would then help the team determine the weighting of each dimension towards the overall data quality score.
+Some of the dimensions were combined or removed.
+For example, granularity and interpretability were removed and interpretability was combined with usability.
+The team then used the ranking weighting method, Sum and Reciprocal, and obtain the following weights for each dimension alongside the type of questions asked for each dimension [@citeCatalogue; @citeCatalogueSteps]:
 
-The score for each dataset is recalculated every week [@citeCatalogue]
+![Weighting](other/images/weighting_of_dimensions.png)
+
+For more technical details about how each dimension is calculated for a dataset and the final data quality score, it can be through the following link: [https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb](https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb) [@citeDQSNotebook].
+However, it is worth noting that each dimension are calculated based on data obtained on the metadata information and the dataset itself.
+If the data quality score of a dataset has a normalized score of less than 60%, they are given the grade "Bronze", if the normalized score is between 60% and 80%, it is given the grade "Silver", and finally if the normalized score is over 80%, tit is given the grade "Gold".
+The score for each dataset in the portal is recalculated every week by the team [@citeCatalogue].
 
@@
 
-\newpage
-
 # References

### 2025-05-15 paper/paper.qmd
Finished data section

@@
 In the results section (@sec-results) we looked at the data more in-depth through graphs.
+With the discussion section (@sec-discussion), we will provide an overview of what was done in our results, discussing our results and its real-world implications, and indicate areas of improvements for our analysis and directions for future works.
 Finally, the appendix (@sec-appendix) includes acknowledgements and any additional information related to the paper.
@@
 In order to create the "Data Quality Score", they assembled the Data Quality Working Group, which consist of a diverse group of people from consumers of datasets to people who create datasets [@citeCatalogue; @citeCatalogueSteps].
+
 Open Data Toronto first reviewed various literature such as academic papers and industry white papers to compile 15 dimensions used to measure quality [@citeCatalogue; @citeCatalogueSteps].
 The dimensions selected were as follows: Interpretability ("How easy it is to understand the data?"), Usability ("How easy is it to work with the data?"), Metadata ("Is the data well described?"), Freshness ("How close to creation time is the data published?"), Granularity ("How atomic is the data?"), Completeness ("How much data is missing?"), and Accessibility ("Is the data easy to access?") [@citeCatalogue; @citeCatalogueSteps].
+
 From there the Data Quality Working Group were surveyed and asked to rank the importance of each dimension when it comes to assessing data quality where 1 represents the most important and 7 being the least important.
@@
 
-![Weighting](other/images/weighting_of_dimensions.png)
+![Weighting](../other/images/weighting_of_dimensions.png)
 
-For more technical details about how each dimension is calculated for a dataset and the final data quality score, it can be through the following link: [https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb](https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb) [@citeDQSNotebook].
+For more technical details about how each dimension is calculated for a dataset and the final data quality score, it can be through the following link:[https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb](https://github.com/open-data-toronto/framework-data-quality/blob/master/data_quality_score.ipynb) [@citeDQSNotebook].
 However, it is worth noting that each dimension are calculated based on data obtained on the metadata information and the dataset itself.
@@
 
-Our variables of interest are "accessibility", "completeness", "freshness", "metadata", "usability", and "grade".
+Our variables of interest that we used in our analysis are the following: "accessibility", "completeness", "freshness", "metadata", "usability", and "grade".
+"Accessiblity" is a score from 0 to 1 that indicates the degree that the dataset can be access or not through the Open Data Toronto API, keywords or tags, and automated data pipelines with 1 being that it can not be accessed with the various methods noted and 0 if not at all.
+"Completeness" is a score from 0 to 1 that indicates how much of the data is missing with 1 being that there is no missing data and 0 being that all data fields are empty or missing.
+The "freshness" variable is a score from 0 to 1 that indicates how up-to-date the data where a shorter time duration between the recent refresh date and time and the previous refresh date and time before that gives a higher score.
+The "metadata" variable indicates how complete the following metadata fields are for a dataset from a scale of 0 to 1: Description, Limitations, Topics, Contact Email.
+The more metadata fields compelted, the higher the metadata score.
+The "usability" variable is a score from 0 to 1 indicating how easy it would be to use the dataset and this is determine by the proportion of meaningful column names or in other words column names with English words in it.
 
@@
 
+# Discussion {#sec-discussion}
+In @sec-results, we
+
+## Point 1
+
+## Point 2
+
+## Areas of improvement
+
+## Next steps
+
+\newpage
+
 \appendix

### 2025-05-15 paper/paper.qmd
Finished introduction and cleaned up data section

@@
 
-In the analysis of our paper, we looked at the following questions: What is the quality of the datasets on the Open Data Toronto portal?
-What are the features of different types of datasets on Open Data Toronto?
+In 2024, a story collaboration between the Investigative Journalism Foundation (IJF) and CBC published in the CBC reported that the risk of death and injuries from fires in lower-income Toronto wards was higher compared to higher-income Toronto wards [@citeCityFire].
+This story as well as other stories published in the news rely on data from Open Data Toronto to bring their stories to life [@citeCityFire].
+Open Data Toronto serves as a hub for all types of data related to Toronto from crime data to information about shelters across the city [@citeODTabout].
+Open Data Toronto has not only been used in the news but also in civic spaces when establishing city policies [@citeODTabout].
+Given the importance of the hub, it raises concerns about the quality of Open Data Toronto's data catalogue and the following question: What is the quality of the datasets like on the Open Data Toronto portal?
+
+In this paper, we analyzed data provided by Open Data Toronto on the data quality grade of the datasets in their catalogue and their characteristics like "accessibility", "completeness", "freshness", "metadata", and "usability".
+Farrow analyzed the data quality scores of the different datasets on the Open Data Toronto portal and found that the metadata and freshness scores were poor overall [@citeFarrow].
+However there has been no analysis done on the scores and how they compare for different graded datasets as of 2025.
+In our findings, all datasets in Open Data Toronto portal are accessible and a majority of them having minimial missing data and are usable.
+It also showed that 56% of their datasets are graded bronze with bronze-graded datasets being less likely to be updated and have completed metadata fields.
+These bronze-graded datasets contribute the poor metadata and freshness scores we saw.
+These findings can inform Open Data Toronto about specific graded datasets that should be given more attention and the specific qualities of them such as metadata completion, which impacts the quality of the data used in the media and in policymaking.
+These findings can also be informative for users of the portal in order to understand what goes behind the grade of datasets on the portal and what they mean.
 
@@
 ## Measurement
-The Information & Technology department at Open Data Toronto collected datasets on their portal using the CKAN Datastore API
-[@citeCatalogue; @citeCatalogueSteps].
 Open Data Toronto uses a metric called the "Data Quality Score" in order to give each dataset in their catalogue a grade indicating "bronze", "silver", or "gold", which can be seen through on the webpage for each dataset on Open Data Toronto portal website.
@@
 
-![Weighting](../other/images/weighting_of_dimensions.png)
+![Weight of the selected dimensions for data quality assessment and the metrics for each dimension](../other/images/weighting_of_dimensions.png)
 
@@
 If the data quality score of a dataset has a normalized score of less than 60%, they are given the grade "Bronze", if the normalized score is between 60% and 80%, it is given the grade "Silver", and finally if the normalized score is over 80%, tit is given the grade "Gold".
+The Information & Technology department at Open Data Toronto collected datasets on their portal using the CKAN Datastore API in order to give each of them a data quality score and grade [@citeCatalogue; @citeCatalogueSteps].
 The score for each dataset in the portal is recalculated every week by the team [@citeCatalogue].
@@
 The "usability" variable is a score from 0 to 1 indicating how easy it would be to use the dataset and this is determine by the proportion of meaningful column names or in other words column names with English words in it.
+However a limitation of the "usability" variables is that it does not consider datasets that are in different languages other than English.
 

### 2025-05-15 paper/paper.qmd
Cleaned up data section and added areas of improvement for discussion section

@@
 From there the Data Quality Working Group were surveyed and asked to rank the importance of each dimension when it comes to assessing data quality where 1 represents the most important and 7 being the least important.
-The results from the survey would then help the team determine the weighting of each dimension towards the overall data quality score.
+The results from the survey would then help the team determine the weight of each dimension towards the overall data quality score.
 Some of the dimensions were combined or removed.
 For example, granularity and interpretability were removed and interpretability was combined with usability.
-The team then used the ranking weighting method, Sum and Reciprocal, and obtain the following weights for each dimension alongside the type of questions asked for each dimension [@citeCatalogue; @citeCatalogueSteps]:
+The team then used the ranking weighing method, Sum and Reciprocal, and obtain the following weights for each dimension alongside the type of questions asked for each dimension [@citeCatalogue; @citeCatalogueSteps]:
 
-![Weight of the selected dimensions for data quality assessment and the metrics for each dimension](../other/images/weighting_of_dimensions.png)
+![Weight of the selected dimensions for data quality assessment and the metrics for each dimension](../other/images/weight.png)
 
@@
 # Discussion {#sec-discussion}
-In @sec-results, we
+In @sec-results, we looked at the data quality of 39,580 datasets on the Open Data Toronto
 
-## Point 1
+## Majority of Datasets are graded "Bronze"
 
-## Point 2
+## Bronze-graded datasets are less likely to update and have missing metadata
 
 ## Areas of improvement 
+As mentioned in @sec-data, datasets have a higher usability score if their column names contains more English words.
+However, this criteria does not consider datasets that could be still useful but are not in English.
+Another limitation of our analysis is that the weight of the different dimensions in the data that goes towards the grade of a dataset is subjective in nature since the weighing is based on survey data that had people rank the perceived importance of each dimension.
 

### 2025-05-15 paper/paper.qmd
Fixed image link in paper
Finished paper

@@
 
-![Weight of the selected dimensions for data quality assessment and the metrics for each dimension](../other/images/weight.png)
+![Weight of the selected dimensions for data quality assessment and the metrics for each dimension](../other/images/weight_of_dimensions.png)
 
@@
 # Discussion {#sec-discussion}
-In @sec-results, we looked at the data quality of 39,580 datasets on the Open Data Toronto
+In @sec-results, we looked at the data quality of 39,580 datasets on the Open Data Toronto as of May 13, 2025 and the different characteristics of the datasets.
+We found that our analysis was consistent with what @citeFarrow found where the metadata and freshness scores of datasets overall was poor but also we found that the low scores were contributed from the bronze-graded datasets.
 
 ## Majority of Datasets are graded "Bronze" 
+We saw with that with @fig-grade-accessibility-bar that 56% of datasets in the Open Data Toronto portal are graded "Bronze".
+This indicates that 56% of datasets had a data quality score of less than 60%.
+This also raises concerns regarding the quality of the datasets used in news report for example as well as bring awareness of the quality of datasets currently on the portal.
+Fortunately based on what @citeFarrow found in comparison to our results in 2025, there has a decrease in bronze-graded datasets since 2021 from 78% to 56%.
 
 ## Bronze-graded datasets are less likely to update and have missing metadata 
+Our results from @fig-metadata-distribution and @fig-freshness-distribution shows that bronze-graded datasets have low metadata and freshness scores close to 0 or 0%.
+This indicates that the bronze-graded datasets contributes to the low metadata and freshness score seen of Open Data Toronto's entire data catalogue.
+As noted by IBM, Metadata plays an important role in "data governance and data management" [@citeMetadataImportance].
+This means that the lack of metadata for a dataset decreases the experience for users and organization of using the datasets leading to potentially consequences due to issues such as the lack of information of the dataset's limitations and the lack of information about the dataset author's and their contact information.
 
@@
 ## Next steps
+Results from our analysis can be used help the Open Data Toronto team figure out the qualities of bronze-graded datasets that leads to their low scores and also be insight for users of the datasets from Open Data Toronto about what goes behind the grade of the datasets.
+Future works regarding Open Data Toronto's catalogue can look into the data quality score of the datasets from different divisions.
 
@@
 
-We would like to thank @tellingstories for providing assistance with the code used to produce the graphs in this paper.
+We would like to thank @tellingstories for providing assistance with the code used to produce the graphs in this paper. We would also like to thank the team at the IJF for their feedback.
 

### 2025-05-15 paper/paper.qmd
Updated abstract

@@
 date-format: long
-abstract: "As one of the central hubs for Toronto-related data, we conducted analysis on the data quality of Open Data Toronto's catalogue.
-We found that despite Open Data Toronto's extensive dataset catalogue being accessible and having minimal missing data, 56% of their datasets are graded bronze and bronze-graded datasets are less likely to be updated and have completed metadata fields.
-These findings can help raise awareness to Open Data Toronto whose datasets play an important role in news reporting and policymaking and anyone who is interested using datasets from Open Data Toronto's catalogue understand what goes behind the grade given to datasets."
+abstract: "As one of the central hubs for Toronto-related data, we analyzed the data quality of Open Data Toronto's catalogue.
+Despite Open Data Toronto's extensive dataset catalogue being accessible and having minimal missing data, 56% of their datasets are graded bronze and bronze-graded datasets are less likely to be updated and have completed metadata fields.
+These findings can help raise awareness to Open Data Toronto whose datasets play an important role in news reporting and policymaking, and also inform anyone interested in using datasets from Open Data Toronto's catalogue about what goes behind the grade given to datasets."
 format:

### 2025-05-15 paper/paper.qmd
Updated intro
Typo fixed

@@
 Farrow analyzed the data quality scores of the different datasets on the Open Data Toronto portal and found that the metadata and freshness scores were poor overall [@citeFarrow].
-However there has been no analysis done on the scores and how they compare for different graded datasets as of 2025.
+However there has been no analysis done on the characteristics and how they compare for different graded datasets as of 2025.
 In our findings, all datasets in Open Data Toronto portal are accessible and a majority of them having minimial missing data and are usable.
@@
 However, it is worth noting that each dimension are calculated based on data obtained on the metadata information and the dataset itself.
-If the data quality score of a dataset has a normalized score of less than 60%, they are given the grade "Bronze", if the normalized score is between 60% and 80%, it is given the grade "Silver", and finally if the normalized score is over 80%, tit is given the grade "Gold".
+If the data quality score of a dataset has a normalized score of less than 60%, they are given the grade "Bronze", if the normalized score is between 60% and 80%, it is given the grade "Silver", and finally if the normalized score is over 80%, it is given the grade "Gold".
 The Information & Technology department at Open Data Toronto collected datasets on their portal using the CKAN Datastore API in order to give each of them a data quality score and grade [@citeCatalogue; @citeCatalogueSteps].

### 2025-05-16 paper/paper.qmd
Made fixes

@@
 However there has been no analysis done on the characteristics and how they compare for different graded datasets as of 2025.
-In our findings, all datasets in Open Data Toronto portal are accessible and a majority of them having minimial missing data and are usable.
+In our findings, all datasets in Open Data Toronto portal are accessible and a majority of them having minimal missing data and are usable.
 It also showed that 56% of their datasets are graded bronze with bronze-graded datasets being less likely to be updated and have completed metadata fields.
@@
 Our variables of interest that we used in our analysis are the following: "accessibility", "completeness", "freshness", "metadata", "usability", and "grade".
-"Accessiblity" is a score from 0 to 1 that indicates the degree that the dataset can be access or not through the Open Data Toronto API, keywords or tags, and automated data pipelines with 1 being that it can not be accessed with the various methods noted and 0 if not at all.
+"Accessibility" is a score from 0 to 1 that indicates the degree that the dataset can be access or not through the Open Data Toronto API, keywords or tags, and automated data pipelines with 1 being that it can not be accessed with the various methods noted and 0 if not at all.
 "Completeness" is a score from 0 to 1 that indicates how much of the data is missing with 1 being that there is no missing data and 0 being that all data fields are empty or missing.
@@
 The "metadata" variable indicates how complete the following metadata fields are for a dataset from a scale of 0 to 1: Description, Limitations, Topics, Contact Email.
-The more metadata fields compelted, the higher the metadata score.
+The more metadata fields completed, the higher the metadata score.
 The "usability" variable is a score from 0 to 1 indicating how easy it would be to use the dataset and this is determine by the proportion of meaningful column names or in other words column names with English words in it.
